\documentclass[12pt]{article}
\usepackage{parskip}


\title{How to use the schedule file}
\author{Andrew MacDonald}
\date{Nov 2012}

\begin{document}

\maketitle


This is a quick guide to the precipitation experiment schedule
file. Files will be named FIELDSITEschedule.csv, where FIELDSITE will
be replaced with the name of your country or site
(e.g. Cardososchedule.csv, FrenchGuianaschedule.cv) The description
follows every column name:


\textbf{trt.name} = treatment name, following the naming convention
suggested by Diane.  Treatments are named after the two parameters of
the negative binomial distribution (\emph{mu} and \emph{k}).
Names are in the form ``muXkY'', where X and Y are numbers which
multiply \emph{mu} and \emph{k}, respectively, in different
treatments.

\textbf{intended.mu} = the actual parameter values used in this field
site. The ``mu1k1'' treatment was calculated from the original data,
the others were derived from mu1k1 by multiplying those values of
\emph{mu} and \emph{k} by the range of factors agreed upon by the
working group (\emph{mu}: 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.5, 2, 2.5, 3
and \emph{k}: 0.5, 1, 2)

\textbf{intended.k} = as above, but for \emph{k}

\textbf{temporal.block} = after intended.mu and intended.k were
calculated, the treatments were divided into three random groups;
these groups begin on different days (the 1st, 2nd and 3rd days of the
experiment).  The grouping was tested with an ANOVA to confirm that
there is no significant difference in intended.mu or intended.k.

\textbf{1-67} = these columns represent ``days from beginning of
experiment''. They do \textbf{not} represent calendar dates.  Each
column gives the amount of water \textbf{in mm} that should be added
to each plant.  Remember to multiply these by the catchment area of
your bromeliad species, \emph{and} by the correction factor for the
canopy, as described by Diane and her team in Costa Rica.  Therefore,
treatments are applied to bromeliads over a period of \textbf{68
  days}, not including the preparation of bromeliads beforehand, the
calculation of correction factors, and the final destructive sampling
at the end.  Plants receive no water on days marked both NA and 0.
(NAs represent days added by the experimental design, for example the
temporal blocks, in contrast to 0s which are predicted by our model of
rainfall).  Additionally, plants are not watered on days marked
``sample'' or ``insects''.  These represent days set aside for the
collection of protist samples (\emph{sample}) and destructive sampling for
macroinvertebrates (\emph{insects}).

Looking along each row (rather than down each column), the pattern of
treatments for each bromeliad is as follows:

\begin{itemize}
\item 0--2 days delay, depending on the temporal.block (marked NA)*
\item 1 day median water of the mu1k1 treatment
\item \emph{sample} the protist community (\emph{no water} on this day)
\item 60 days of treatment watering
\item 1 day treatment median
\item \emph{sample} the protist community again (\emph{no water} on
  this day)
\item one NA day with no watering or sampling
\item destructively sample the bromeliad for all macroinvertebrates
  (\emph{insects})
\item 0--2 days delay, depending on temporal.block*
\end{itemize}

*The number of these days always sums to 2 (i.e. they are either at
the beginning or end of the experiment, depending on temporal.block)

Thus the full treatment period takes \textbf{68 days}.

Note that the NA days at the end (i.e. on days 67 or 68) are simply
included to keep the schedule looking ``square''.

\end{document}
